---
title: "cocktail"
author: "Hesam Parsa"
date: "June 12, 2016"
output: html_document
---



```{r }

library(jsonlite)

by_id <- fromJSON("http://www.thecocktaildb.com/api/json/v1/1/lookup.php?i=15112")
setwd("~/Desktop/Bootcamp/three_partB")
```

```{r}
database=data.frame()
 for(i in 10000:20000) {
  link="http://www.thecocktaildb.com/api/json/v1/1/lookup.php?i="
   z=paste(link,i,sep="")
   data=as.data.frame(fromJSON(z))
   database=rbind(database,data)
   print(i)
   write.csv(database, file="database.csv" )
 }
```

```{r}
library(dplyr)
data_base <- read.csv("database.csv", header=T, na.strings=c(""))
head(data_base)

db <- data_base %>%
  select(cocktail.name = drinks.strDrink,
         drinks.strIngredient1:drinks.strIngredient7,
         drinks.strMeasure1:drinks.strMeasure7)
  
total_na <-sum(is.na(db))
total_na

db[ db =="" ] = NA
db[ db ==" " ] = NA
db[ db =="/t" ] = NA
db[ db =="\n" ] = NA

# for this to work convert to character
db <- as.data.frame(lapply(db, as.character), stringsAsFactors=FALSE)

db[is.na(db)] <- 0

total_na <- sum(is.na(db))
total_na

```

```{r}
install.packages("tidyr")
library(tidyr)
# lets select name and ingridients. 
db_ingridient <- db %>%
 select(cocktail.name:drinks.strIngredient7)

db_ingridient %>% head()

# make ingridient tidy

db_ing_tidy <- db_ingridient %>%
gather(ingridients_number, ingridient, - cocktail.name)

db_ing_tidy %>% head()
db_ing <-db_ing_tidy

# Of course the above steps can be combined in the piping operator.

# lets select name and ingridients. 
db_measure <- db %>%
 select(cocktail.name, drinks.strMeasure1:drinks.strMeasure7)

db_measure %>% head()

# make ingridient tidy

db_m_tidy <- db_measure %>%
gather(ingridients_number, measure, - cocktail.name)

db_m_tidy %>% head()

db_m <- db_m_tidy

db_join <- db_ing %>%
  bind_cols(db_m)

db_tidyy <- db_tidyy[, -3]

db_tidy <- db_tidyy %>%
  distinct() %>%
  filter (ingridient !=0 | measure !=0)



db_tidy$measure[db_tidy$measure == 0]  <- 5

db_top10 <- db_tidy 

db_top10_grp <- db_top10 %>% 
 group_by(ingridient) %>%
 summarise(N = n()) %>%
 arrange(desc(N))

db_top10_grp

``` {r}
library(stringr)
# for mac users. remove white spaces. 
db_tidy <-mutate(db_tidy, 
               amount = str_trim(db_tidy$measure, side = "both"))




# separate numbers into its own column.

# We will create a new column in each step to see the reduction in anomalities.

# create a new column with numbers with punctuation. 
db_tidyz <- db_tidy %>%
mutate(number=gsub("[^[:digit:],^[:punct:] ]", "", amount)) %>% #head()
  
# Trim both sides from whitespaces
mutate(numbe=str_trim(number,side="both"))%>% 
  
# replace double and triple and so on whitspaces with single whitespace
 
mutate(numb=gsub("\\s+"," ",numbe))%>% #head()
  
# Select all numbers that matching the following four patterns
mutate(num=str_extract(numb,"[:digit:]+\\s+[:digit:]+[:punct:]+[:digit:]|
                      [:digit:]+[:punct:]+[:digit:]+|
                      [:digit:]+|
                      [:digit:]+[:punct:]+[:digit:]+[:digit:]+"))%>% #head()

# Replace all whitespaces with + sign
  mutate(nu=str_replace_all(num," ","+")) %>% #head()

# Separate everything that has "-" into two columns
separate(nu,c("nu","nuu"), sep="-") %>%  # head()

# Select text only (with white spaces)   
mutate(unit=gsub("[[:digit:]|[:punct:]]","",amount))%>% # head()
  
# Trim both sides from whitespaces
mutate(unit=str_trim(unit,side="both")) %>%
  
#replace double and triple and so on whitspaces with single whitespace
mutate(unit=gsub("\\s+"," ",unit))



# set NAs to 1
db_tidyz[is.na(db_tidyz)]<-1



#Check unique values for number, numbe, num and unit
unique(db_tidyz$number)
unique(db_tidyz$numbe)
unique(db_tidyz$numb)
unique(db_tidyz$num)
unique(db_tidyz$nu)
unique(db_tidyz$unit)

{r ingrindient_to_ml}
# Select only cocktail name, ingredient, nu number and unit column
db_tidyx <- db_tidyz %>%
   select(cocktail.name,ingridient,nu, unit)

db_tidyx %>% head()

# Replace units with proper conversion to mL
db_tidyx <- db_tidyx %>%
   mutate(unit=str_replace(unit,"ozjamaican","oz")) %>%
   mutate(unit=str_replace(unit,"oz","29.5")) %>%
   mutate(unit=str_replace(unit,"shot","29.5")) %>%
   mutate(unit=str_replace(unit,"jigger","44.5")) %>%
   mutate(unit=str_replace(unit,"cup","257")) %>%
   mutate(unit=str_replace(unit,"tblsp","11.1")) %>%
   mutate(unit=str_replace(unit,"tsp","3.7")) %>%
   mutate(unit=str_replace(unit,"ts p","3.7")) %>%
   mutate(unit=str_replace(unit,"teaspoon","3.7")) %>%
   mutate(unit=str_replace(unit,"cl","10")) %>%
   mutate(unit=str_replace(unit,"dl","100")) %>%
   mutate(unit=str_replace(unit,"litre","1000")) %>%
   mutate(unit=str_replace(unit,"liter","1000")) %>%
   mutate(unit=str_replace(unit,"dash","0.9")) %>%
   mutate(unit=str_replace(unit,"splash","3.7")) %>%
   mutate(unit=str_replace(unit,"twist","15")) %>%
   mutate(unit=str_replace(unit,"twistof","15")) %>%
   mutate(unit=str_replace(unit,"can","355")) %>%
   mutate(unit=str_replace(unit,"cube","12")) %>%
  mutate(unit=str_replace(unit,"part","29.5")) %>%
   mutate(unit=str_replace(unit,"pint","473")) %>%
   mutate(unit=str_replace(unit,"glass","473"))

# Check if missing something like glass or a pint add it to the code on top
unique(db_tidyx$unit)
```

```{r dealing_with_fractions}


db_tidyc<-db_tidyx %>%
   # Select all numbers that matching the following patterns
   mutate(unit=str_extract(unit,"[:digit:]+[:digit:]+[:punct:]+[:digit:]|[:digit:]+[:punct:]+[:digit:]+|[:digit:]+|[:digit:]+[:punct:]+[:digit:]+[:digit:]+"))%>%   #head()
   
   # Convert unit column to numeric
   mutate(unit=as.numeric(unit))%>%  
   
   # replace comma , with dot .
   mutate(nu=gsub(",",".",nu)) %>%  
   # add left and right brackets for nu colum for proper parsing and evaluation
   mutate(nu=sub("^", "(",nu)) %>%   #head()
   mutate(nu=sub("$", ")",nu))  #%>%  head()

# set NAs to 1
db_tidyc[is.na(db_tidyc)]<-1
      
# parse and evaluate nu
m2 <- sapply(db_tidyc$nu ,function(x) eval(parse(text=x)))

str(m2)

# add m2 as a new column 
db_tidyc$nup<-m2

# compare nu and nup for sanity check 
head(db_tidyc,10)

# looks good

# multiply nup and unit for the measure
db_tidyc<- db_tidyc %>%
   mutate(measure=nup*unit)

db_tidyc %>% head(10)

#  Select only cocktail name, ingridient and measure
db_clean_tidy <- db_tidyc %>%
   select(cocktail.name,ingridient,measure)

db_clean_tidy %>% head(10)
```

```{r spread}
# Use tidyr to spread the data. This is the opposite of gather. However for this case it does not work automatically, creates duplicated rows.  
# Hadley's save the spread with unique identifiers:

db_clean_tidy$row <- 1:nrow(db_clean_tidy)

# Let's spread it!
db_spread <- db_clean_tidy %>%
   spread(ingridient, measure)

# Remove the row column
db_spread<- db_spread %>%
   select(-row)

# Replace NAs with 0s
db_spread[is.na(db_spread)]<-0

# Group by cocktail name
db_spread_comb <- db_spread %>%
   group_by(cocktail.name) %>%
   summarise_each(funs(sum))
# Check ingridients for a random cocktail
i=37
colnames(db_spread_comb)[which(db_spread_comb[i,]!=0)]

"salt" %in% colnames(db_spread_comb)

# Spread is ready! Let's save it!
# write.csv(db_spread,"db_spreaded.csv")
```

```{r}
# Perform K-means clustering with 10 possible groups, add clusters as additional column 
dtf <- db_spread_comb[, -1]
km.out <- kmeans (dtf, 10, nstart=20, iter.max=50)
dclust<-data.frame(km.out$cluster,db_spread_comb$cocktail.name)

# Save clustered data 
write.csv(dclust,"clusteredata.csv")
```

```{}
# Trim Ingredient Term Matrix, remove cocktail names, turn to numeric
dtf<-db_spread_comb[,-1]
dtf <- sapply(dtf, as.numeric)

# Replace NAs with 0 and turn into dataframe, explore structure

dtf[is.na(dtf)]<-0

dtf<-as.data.frame(dtf)
str(dtf)

# Calculate Ingredients Correlation Matrix, set NAs to 0
xcorr<-cor(dtf)
xcorr[is.na(xcorr)]<-0

# Explore Ingredients Correlation Matrix for values above 65%

cor_mat<-xcorr
most_cor <- which(abs(cor_mat) > 0.75 & cor_mat != 1,
                 arr.ind=TRUE)

rownames(cor_mat)[7]
```


```{}
angle <- function(x,y){
 dot.prod <- x%*%y 
 norm.x <- norm(x,type="2")
 norm.y <- norm(y,type="2")
 theta <- acos(dot.prod / (norm.x * norm.y))
 as.numeric(theta/3.14*180)
}


# Prepare ingredient term matrix
dtf <- sapply(dtf, as.numeric)

# Create dummy inner product matrix
sumi<-matrix(nrow=nrow(dtf),ncol=nrow(dtf))

# Test if angle function works, answer is in degrees
angle(dtf[100,],dtf[220,])

# Apply angle function on ingredient term matrix (without cocktail names), row by row

for (i in 1:nrow(dtf)){
  for (j in 1:nrow(dtf)){
   sumi[i,j]<-angle(dtf[i,],dtf[j,])
  # print(i)
  # print(j)
   }
}
# Replace NAs with 0
sumi[is.na(sumi)]<-0

# Get histogram
hist(sumi)
# Convert to dataframe set rows and columns names to the names of cocktails 
sumidf<-as.data.frame(sumi)
colnames(sumidf)<-db_spread_comb$cocktail.name
rownames(sumidf)<-colnames(sumidf)

# Replace NAs with 0s
sumidf[is.na(sumidf)]<-0

# Save as db_innerproduct_matrix.csv
#write.csv(sumi,"cocktailz/db_innerproduct_matrix.csv")

[5:12]  
## Explore inner products matrix

```{r Explore_Inner_Products, eval=FALSE}
# Load Inner Products Matrix, set rownames the same as column names
#dfsumi<-read.csv("db_innerproduct_matrix.csv", header = TRUE)
dfsumi<-dfsumi[,-1] # take out row number

colnames(dfsumi) <- db_spread_comb$cocktail.name[1:3156]
rownames(dfsumi)<-colnames(dfsumi)

# set NAs to zero
dfsumi[is.na(dfsumi)] <- 0

# Choose Margarita
x<-dfsumi %>%
  select(Margarita)
rownames(x)<-colnames(dfsumi)
x<-sapply(x, as.numeric)

y<-colnames(dfsumi)
hist(x[x<70])

# Perform unsupervised kmeans clustering on inner products matrix
km.out.ip <- kmeans(dfsumi, 10, nstart =20,iter.max=50)

# previous cluster
ip.x <-as.data.frame(km.out.ip$cluster)

ip.x <- bind_cols(ip.x, as.data.frame(rownames(ip.x)))
```


